Negative and Fractional Types

### 2012-06-29 rational_ext.tex
meadows

Without the important side condition on division by zero, we would
have {{1<->1/0*0}} which implies {{1 <-> 0}} consequently collapsing
the entire structure. In this context, an important area for future
investigation is to consider alternatives to fields as a model for
fractionals and negatives -- particularly algebraic structures that
are ``zero totalized fields'' such as meadows
\cite{Bergstra:2009:MES:1507774.1508000}.

@@

@@

Charles Pinter in his book on Abstract Algebra talks about how
negatives were used by \emph{logisticians} of a 1000 years ago used
negative numbers as a ``useful fiction''.  The concept of negative
numbers were ``discovered''much later than fractionals. Like Joan Baez
says ``half an apple is easier to undertand than a negative apple''.

Negative types should correspond to sets with negative arity. But what
does that mean? Various approaches have been proposed -- Loeb uses the
notion of multiplicity functions to express negative sized sets
[CITE]. These notions have been extended to Euler characteristic of
categories by Leinster, Joyal and Schanuel -- negative values are
explained sometimes as eqaulities relating otherwise indepedant
objects -- Baez tells a story about ``island'' and positive and
``bridges connecting islands'' as negatives (since the effectively
reduce trhe number of disconnected elements.)

We use the use the syntax {{-b}} for negative types and {{-v}} for
negative values and the axiom {{0 <-> (-b)+b}} as the basis for
negatives. In {{Pi}}/{{PiEE*}}, values flowed from left-to-right in
circuits. Negative types are negative in the sense that they flow in
the oppostite direction.

@@

\begin{definition}[Size of a type]
The defintion of size is extended to include negative types.
{{[^-b^] = -[^b^]}}
\end{definition}

>>>>

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Zero Totalized Fields}

Without the important side condition on division by zero, we would
have {{1<->1/0*0}} which implies {{1 <-> 0}} consequently collapsing
the entire structure.

While fields require the ``no division by zero'' constraint, alternate
algebraic structures such as meadows
\cite{Bergstra:2009:MES:1507774.1508000} have been studied which admit
division and negative without this constraint. Such structures are
called ``zero totalized fields'' and define multiplicative inverse of
{{0}}. The axiomatic presentation of meadows involve removing the
combinators corresponding the axiom {{0*b<->0}} (i.e. {{dist0}} and
{{factor0}}) and {{1<->1/b*b}} (i.e. {{eta*}} and {{eps*}}). Instead
meadows include axioms \emph{reflexivity}, {{refl}} and {{refr}}, and
\emph{restricted inverse}, {{ril}} and {{rir}}, which are:
%subcode{opsem} include main
% refl :& 1/1/b <-> b :& refr
% ril :& b*(b*1/b) <-> b :& rir

\noindent
Combinators {{refl}} and {{refr}} are primitive and can be defined as:
%subcode{opsem} include main
% refl 1/(1/v) &|-->& v
% refr v &|-->& 1/(1/v)

\noindent
Combinators {{ril}} and {{rir}} are defined as machine transitions
similar to {{eta*}}/{{eps*}}:
%subcode{opsem} include main
% <ril, (v,(v',1/v')), C> &|-->& {[ril, v, C]}
% <rir, v, C> &|-->& forall v':b.{[rir, (v, (v',1/v')), C]}

While not as commonly used as fields, zero totalization is useful not
just because we are spared from continuously checking the division
constraint, but also because many other structures such as the
category of vectors spaces do admit a dual to the 0 vector
space. However their categorization is not well developed and we
however lose the nice correspondence that {{eta*}}/{{eps*}} have with
compact closed structures.

@@

Charles Pinter in his book on Abstract Algebra talks about how
negatives were used by \emph{logisticians} of a 1000 years ago used
negative numbers as a ``useful fiction''.  The concept of negative
numbers were ``discovered''much later than fractionals. Like Joan Baez
says ``half an apple is easier to understand than a negative apple''.

Negative types should correspond to sets with negative arity. But what
does that mean? Various approaches have been proposed -- Loeb uses the
notion of multiplicity functions to express negative sized sets
[CITE]. These notions have been extended to Euler characteristic of
categories by Leinster, Joyal and Schanuel -- negative values are
explained sometimes as equalities relating otherwise independent
objects -- Baez tells a story about ``island'' and positive and
``bridges connecting islands'' as negatives (since the effectively
reduce the number of disconnected elements.)

We use the use the syntax {{-b}} for negative types and {{-v}} for
negative values and the axiom {{0 <-> (-b)+b}} as the basis for
negatives. In {{Pi}}/{{PiEE*}}, values flowed from left-to-right in
circuits. Negative types are negative in the sense that they flow in
the opposite direction.

@@

\begin{definition}[Size of a type]
The definition of size is extended to include negative types.
{{[^-b^] = -[^b^]}}
\end{definition}
